Perceptions of Corruption in Nigeria

### 2024-12-01 paper/paper.qmd
packages referencing

---
title: "Perceptions of Corruption in Nigeria"
subtitle: "Rural Nigerians and Less Educated  Groups Report Higher Corruption"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

#install.packages("modelsummary")
#install.packages("rstanarm")
#install.packages("dplyr")
#install.packages("here")
#install.packages("arrow")
library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)

library(patchwork)

@@

We use the statistical programming language R [@citeR]....

@@

For this analysis, i use the Afrobarometer Rouund 9, Nigeria, 2023
survey. Afrobarometer conducts surveys that are nationally
representative across African nations to capture public perceptions on
topics like governance, democracy, and economic issues. The Nigerian
Round 9 surveey dataset contains responses from 1,556 people and is
refective of the diverse demographic statistics across gender, region,
and residence type(urban or rural). The data is freely available on the
Afrobarometer website[@afrobarometer2023] and was collected through a
multi-stage, stratified sampling method. This is to ensure
representativeness and enables analysis of different demographics and
socioeconomic indicators.

Our primary measurement task is to capture public opininon towards
economic, political and socioeconomic issues and translate it into
actionable intell. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses.

@@

## Variables of Interest
```{r}
#| label: fig-urban-rural-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: " Urban vs Rural Respondent Distribution. This bar chart compares the number of respondents residing in urban versus rural areas."

@@

```{r}
#| label: fig-corruption-presidency
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Perceived Corruption in the Presidency. This bar chart shows the distribution of respondents' perceptions of corruption in the presidency."

@@

For the analysis, only the substantive responses (0-3) were considered,
in order to help us look into perceptions of corruption within the presidency. 
The high frequency of responses in the "High Corruption" and "Some
Corruption" categories reflects persistent concerns about the integrity
of key officials in Nigeria, aligning with broader findings on
corruption in the region.@fig-corruption-presidency presents the distribution of 
respondents' perceptions of corruption within the Nigerian presidency. Among
these, the category "High corruption" is the most frequently reported,
followed by "Some corruption". "All corrupt" ranks third, with "No
Corruption" being the least chosen category which is no surprise given
our data is in the context of Nigeria. These proportions show
significant public concern about corruption within the presidency,
consistent with broader trends in public concern abut corruption
corruption within the presidency in Nigeria.

```{r}
#| label: fig-age-group-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Age Group Distribution. This bar chart displays the number of respondents in each age group."

@@

```
@fig-age-group-distribution illustrates the distribution of survey respondents across five distinct age groups: 18-24, 25-34, 35-44, 45-54, and 55+. The data reveals that the 25-34 age group is the most represented, with over 500 respondents, followed by the 35-44 age group. The youngest cohort, aged 18-24, comprises approximately 300 respondents, while the older age groups (45-54 and 55+) are less represented in the dataset, with fewer than 200 respondents each. This distribution highlights a concentration of responses from younger demographics, particularly those aged 25-34, which may reflect broader societal trends such as population age structure or survey reach and accessibility.
```{r}
#| label: fig-trust-political-institutions
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Distribution of Trust in Political Institutions. This histogram shows how respondents rated their trust in political institutions, from low to high trust."

library(haven)

# Convert haven_labelled variables to appropriate types
cleaned_data <- cleaned_data %>%
  mutate(
    Q38A = as.numeric(as.character(Q38A)), # Convert to numeric
    Q94 = as.numeric(as.character(Q94)),   # Convert to numeric
    Q101 = as.numeric(as.character(Q101)), # Convert to numeric if needed
    URBRUR = as_factor(URBRUR),            # Convert to factor
    corruption_presidency = as_factor(corruption_presidency), # Convert to factor
    urban_rural = as_factor(urban_rural)   # Convert to factor
  )

# Plot: Trust in Political Institutions
ggplot(cleaned_data, aes(x = Q38A)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(
    title = "Trust in Political Institutions",
    x = "Trust Level (1 = Low, 5 = High)",
    y = "Frequency"
  ) +
  theme_minimal()

@@

```{r}
#| label: fig-demand-democracy
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Distribution of Demand for Democracy. This histogram illustrates respondents' levels of demand for democracy, with higher values reflecting stronger democratic preferences."

# Plot: Demand for Democracy
ggplot(cleaned_data, aes(x = Q94)) +
  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
  labs(
    title = "Demand for Democracy",
    x = "Demand Level (1 = Low, 5 = High)",
    y = "Frequency"
  ) +
  theme_minimal()

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome
variable or update the subheading to be singular.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine
a few into one if they go together naturally.

Background details and diagnostics are included in [Appendix
-@sec-model-details].

@@

The use of these priors provides flexibility for the model while
preventing overfitting. The posterior distributions for the coefficients
and thresholds, along with diagnostic checks for model convergence and
predictive accuracy, can be found in @sec-appendix.

>>>>

---
title: "Perceptions of Corruption in Nigeria"
subtitle: "Rural Nigerians and Less Educated  Groups Report Higher Corruption"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

#install.packages("modelsummary")
#install.packages("rstanarm")
#install.packages("dplyr")
#install.packages("here")
#install.packages("arrow")
#install.packages(tinytable)
library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)
library(patchwork)

@@

@@

For this analysis, i use the Afrobarometer Round 9, Nigeria, 2023
survey. Afrobarometer conducts surveys that are nationally
representative across African nations to capture public perceptions on
topics like governance, democracy, and economic issues. The Nigerian
Round 9 survey dataset contains responses from 1,556 people and shows the diverse demographic statistics across gender, region,
and residence type(urban or rural). The data is freely available on the
Afrobarometer website[@afrobarometer2023] and was collected through a
multi-stage, stratified sampling method. This is to ensure
representativeness and enables analysis of different demographics and
socioeconomic indicators.

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses.

@@

The data for this study was downloaded, cleaned, analyzed, modeled, and visualized using R [@citeR], a statistical programming language. The following packages were used:

- **`modelsummary`** [@modelsummary]: Used to create well-formatted regression tables and summaries.
- **`rstanarm`** [@rstanarm]: Facilitated Bayesian regression modeling using the Stan backend.
- **`dplyr`** [@dplyr]: Streamlined data manipulation, transformation, and summarization processes.
- **`here`** [@here]: Simplified the organization of file paths, improving project reproducibility.
- **`arrow`** [@arrow]: Enabled efficient handling of large datasets with fast reading and writing capabilities.
- **`bayesplot`** [@bayesplot]: Assisted in producing diagnostic checks and posterior predictive visualizations for Bayesian models.
- **`ggplot2`** [@ggplot2]: Provided extensive tools for creating clear and impactful visualizations.
- **`patchwork`** [@patchwork]: Combined multiple plots into cohesive, professional visuals.
- **`knitr`** [@knitr]: Enabled dynamic report generation, integrating code, output, and text seamlessly.
- **`kableExtra`** [@kableExtra]: Enhanced table styling and formatting for clean, professional presentations.
- **`ggalt`** [@ggalt]: Allowed the creation of specialized plots, enhancing the depth of data visualizations.
- **`haven`** [@haven]: Facilitated the import of `.sav` survey data from the Afrobarometer.
- **`broom.mixed`** [@broommixed]: Streamlined the extraction and tidying of results from mixed models for easy analysis and visualization.
- **`afrobarometer`** [@afrobarometer2023]: Served as the primary data source, providing valuable survey data for analysis.
- **_Telling Stories with Data_** [@tellingstories]: This book was referenced for its code and methodologies.

## Cleaned Data Overview 
```{r}
#| label: tbl-cleaned-data-preview
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| tbl-cap: "Preview of the Cleaned Dataset"

#install.packages("knitr")
#install.packages("kableExtra")
library(knitr)
library(kableExtra)

# Create a preview of the cleaned dataset
table_preview <- cleaned_data %>% 
  select(-c(1:6)) %>%
  head(6)

# Generate a well-formatted table using kableExtra
kable(
  table_preview, 
  caption = "Preview of the Cleaned Dataset"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

@tbl-cleaned-data-preview represents the first six cleaned rows of the
dataset, showing the important variables used in the analysis. The
variables were constructed from specific survey questions in the
Afrobarometer Round 9 Nigeria Survey [@afrobarometer2023], which
captures several socioeconomic sentiments within the country.The column
corruption_presidency captures perceptions of corruption within the
office of the presidency, derived from responses to the question, “How
many of the following people do you think are involved in corruption, or
haven’t you heard enough about them to say: the president and officials
in his office?” The variable education reflects respondents' highest
level of education completed, gathered from the question, “What is the
highest level of education you have completed?” The gender variable
indicates the respondent's gender, collected through direct demographic
inquiry. The column age_group categorizes respondents into age brackets
based on the question, “How old are you?” The urban_rural variable
distinguishes between urban and rural respondents, as recorded by
interviewers based on the primary sampling unit (PSU) during data
collection [@survey_manual_2022]. Lastly, region specifies the
geographical region of each respondent, identified using a unique code
for Nigerian states. This table not only demonstrates the diversity of
the dataset but also highlights the breadth of demographic and
governance-related variables included for analysis.

```{r}
#| label: tbl-summary-statistics
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| tbl-cap: "Detailed Summary Statistics of the Cleaned Dataset"

library(dplyr)
library(knitr)
library(kableExtra)

# Ensure the age_group column is a character for correct mapping
cleaned_data <- cleaned_data %>%
  mutate(age_group = as.character(age_group))

# Function to calculate the mode (most frequent category)
calculate_mode <- function(x) {
  x <- na.omit(x)
  unique_x <- unique(x)
  unique_x[which.max(tabulate(match(x, unique_x)))]
}

# Adjust line breaks for proper rendering
line_break <- ifelse(knitr::is_html_output(), "<br>", "\n")

# Prepare the data for the transposed table
transposed_summary <- tibble(
  Statistic = c(
    "Most Frequent Age Group", 
    "Urban Percentage", 
    "Male Percentage", 
    "Corruption Counts", 
    "Education Counts"
  ),
  Value = c(
    calculate_mode(cleaned_data$age_group), 
    paste0(round(mean(cleaned_data$urban_rural == "Urban Residents", na.rm = TRUE) * 100, 2), "%"),
    paste0(round(mean(cleaned_data$gender == "Male", na.rm = TRUE) * 100, 2), "%"),
    paste0(
      "None: 61", line_break,
      "Some: 574", line_break,
      "High: 593", line_break,
      "All: 328"
    ),
    paste0(
      "Primary: 203", line_break,
      "Secondary: 173", line_break,
      "University: 102", line_break,
      "Postgraduate: 07"
    )
  )
)

# Generate a formatted table
kable(
  transposed_summary, 
  col.names = c("Statistic", "Value"),
  caption = "Detailed Summary Statistics of the Cleaned Dataset (Transposed)",
  escape = FALSE # Disables escaping for proper HTML/PDF formatting
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(2, width = "20em") # Adjust width for better multiline display

```
The cleaned dataset used in this study is summarised in @tbl-summary-statistics, which provides key statistics to offer an overview of the dataset's structure and characteristics. The most frequent age group among respondents is identified, representing the predominant demographic in terms of age. The table also highlights the urban and male percentages, which indicate the proportion of urban residents (43.38%) and male respondents (49.68%) in the sample.

Further, the table breaks down the perceptions of corruption into four categories—"None," "Some," "High," and "All"—with counts for each, offering insights into respondents' views on corruption in the presidency. Additionally, the table details the distribution of respondents across education levels, including categories such as "Primary," "Secondary," "University," and "Postgraduate." Together, these summary statistics provide a comprehensive snapshot of the dataset, enabling a clear understanding of its demographic and categorical variables.

## Variables of Interest
```{r}
#| label: fig-urban-rural-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: " Urban vs Rural Respondent Distribution. This bar chart compares the number of respondents residing in urban versus rural areas in Nigeria."

@@

```{r}
#| label: fig-corruption-presidency
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Perceived Corruption in the Presidency. This bar chart shows the distribution of respondents' perceptions of corruption in the presidency in Nigeria."

@@

For the analysis, only the substantive responses (0-3) were considered,
in order to help us look into perceptions of corruption within the
presidency. The high frequency of responses in the "High Corruption" and
"Some Corruption" categories reflects persistent concerns about the
integrity of key officials in Nigeria, aligning with broader findings on
corruption in the region. @fig-corruption-presidency presents the
distribution of respondents' perceptions of corruption within the
Nigerian presidency. Among these, the category "High corruption" is the
most frequently reported, followed by "Some corruption". "All corrupt"
ranks third, with "No Corruption" being the least chosen category which
is no surprise given our data is in the context of Nigeria. These
proportions show significant public concern about corruption within the
presidency, consistent with broader trends in public concern abut
corruption corruption within the presidency in Nigeria.

```{r}
#| label: fig-dumbbell-education
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Distribution of Education Levels in Nigeria"
#install.packages("ggalt")
library(ggalt)

cleaned_data %>%
  count(education) %>%
  ggplot(aes(x = n, y = reorder(education, n))) +
  geom_dumbbell(
    aes(xend = 0, x = n), 
    size = 3, colour = "darkblue", dot_guide = TRUE
  ) +
  theme_minimal() +
  labs(
    title = "Distribution of Education Levels in Nigeria",
    x = "Number of Respondents", y = "Education Level"
  )

@fig-dumbbell-education visually represents the educational attainment
levels of respondents in the Afrobarometer survey for Nigeria. The data
reflects responses to the question: *What is your highest level of
education?*. The chart indicates that the most frequent reported level
of education is "secondary school completed," followed by "primary
school completed" and "no formal schooling". Fewer respondents reported
having attained "post secondary qualifications," "university completed,"
or "Post-graduate" education, shows the disparity in advanced
educational attainment across the population. This distribution shows
that basic education levels are more prevalent among respondents, which
may influence their perceptions and responses in the broader survey.
Including educational attainment is important in our analysis because it
is an important socioeconomic factor.

```{r}
#| label: fig-age-group-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Age Group Distribution. This bar chart displays the number of respondents in each age group in Nigeria."

@@

@@

@fig-age-group-distribution illustrates the distribution of survey
respondents across five distinct age groups: 18-24, 25-34, 35-44, 45-54,
and 55+. The data reveals that the 25-34 age group is the most
represented, with over 500 respondents, followed by the 35-44 age group.
The youngest cohort, aged 18-24, comprises approximately 300
respondents, while the older age groups (45-54 and 55+) are less
represented in the dataset, with fewer than 200 respondents each. This
distribution highlights a concentration of responses from younger
demographics, particularly those aged 25-34, which may reflect broader
societal trends such as population age structure or survey reach and
accessibility.

@@

The use of these priors provides flexibility for the model while
preventing overfitting. The posterior distributions for the coefficients
and thresholds, along with diagnostic checks for model convergence and
predictive accuracy, can be found in the Appendix.
